---
title: "Residual Analysis and Different Models for Electricity Demand"
author: "Sulalitha Bowala"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
rm(list=ls(all=TRUE)) # Remove objects from environment
set.seed(12)
```


## Ontario Data


```{r, message=FALSE}
### Call packages
require("tseries") # load the tseries library
require("zoo")
require("moments")
require("car")
require("MASS")
require("stats")  
require("VGAM")
require("fGarch")
require("quantmod") # To download data alternatively
require("PerformanceAnalytics") # To draw Timeseries graphs
require("scales")
require("PEIP")
require("dplyr")
require("lubridate")
require("tidyverse")
require("gridExtra")
require("gdata")
require("xtable")
require("vioplot")
require("fpp3")
require("readr")
require("tis")
require("tibble")
require("tsibble")
require("stringr")
require("magrittr")
require("fable")
require("Metrics") # rmse()
require("forecast")
```



```{r}
## Import data

Ontario_demand_hourly <- read.csv("Data_2018_2019.csv")
```


```{r}
### Manupulate hourly data to obtain daily data
### + create a new categorical variable (day-type)

## Ontario data

Ontario_demand_hourly$Time <- as.POSIXct(Ontario_demand_hourly$Time, "%Y-%m-%d %H:%M:%S")
Ontario_demand_hourly$Demand <- Ontario_demand_hourly$Demand
Ontario_demand_hourly$Temperature <- Ontario_demand_hourly$Temperature
Ontario_demand_hourly$Precipitation <- Ontario_demand_hourly$Precipitation
Ontario_demand_hourly$CloudCover <- Ontario_demand_hourly$CloudCover
Ontario_demand_hourly$IrradianceSurface <- Ontario_demand_hourly$IrradianceSurface
Ontario_demand_hourly$IrradianceToa <- Ontario_demand_hourly$IrradianceToa
Ontario_demand_hourly$Snowfall <- Ontario_demand_hourly$Snowfall
Ontario_demand_hourly$SnowMass <- Ontario_demand_hourly$SnowMass
Ontario_demand_hourly$Location <- Ontario_demand_hourly$Location
Ontario_demand_hourly <- as_tsibble(Ontario_demand_hourly,index = 'Time',key = 'Location')

Ontario_demand_hourly <- Ontario_demand_hourly %>% 
  mutate(
    Date = date( Time)
  )
Ontario_demand_hourly <- Ontario_demand_hourly %>% 
  mutate(
    Holiday = isHoliday(Ontario_demand_hourly$Date)
  )
Ontario_demand_hourly <- Ontario_demand_hourly %>%
 mutate(Day_Type = case_when(
    Holiday ~ "Holiday",
    wday(Date) %in% 2:6 ~ "Weekday",
    TRUE ~ "Weekend"
  ))
# 
head(Ontario_demand_hourly)
tail(Ontario_demand_hourly)
```


```{r}
## Rename columns

Ontario_demand_hourly<-Ontario_demand_hourly[,c("Time","Precipitation","Temperature","IrradianceSurface","IrradianceToa","Snowfall","SnowMass","CloudCover","AirDensity","Demand","Location","Date","Holiday","Day_Type")]
```




```{r}
## Daily data

# Ontario data
Ontario.Daily <- Ontario_demand_hourly %>%

  index_by(Date = date(Time)) %>%
  summarise(
    Demand = mean(Demand),
    Temperature = mean(Temperature),
    Holiday = any(Holiday),
    Precipitation = mean(Precipitation),
    IrradianceSurface = mean(IrradianceSurface),
    IrradianceToa = mean(IrradianceToa),
    Snowfall = mean(Snowfall),
    SnowMass = mean(SnowMass),
    CloudCover = mean(CloudCover),
    AirDensity = mean(AirDensity),
  ) %>%
  mutate(Day_Type = case_when(
    Holiday ~ "Holiday",
    wday(Date) %in% 2:6 ~ "Weekday",
    TRUE ~ "Weekend"
  ))

Ontario.Daily <- Ontario.Daily %>%
  mutate(Temperature2 = Temperature^2,
         HolidayNew = if_else(Holiday == "TRUE", 1, 0),
         Day_TypeNew = if_else(Day_Type == "Weekday", 1, 0))

head(Ontario.Daily)
tail(Ontario.Daily)
```


whether data: https://www.renewables.ninja/

We believe Temperature is affected by other covariates (Precipitation, IrradianceSurface, IrradianceToa, Snowfall, SnowMass, CloudCover, AirDensity)  available in the data.

Description of covariates:

Air temperature (°C): Air temperature at 2 meters above ground. T2M variable in MERRA-2, converted from K.

Precipitation (mm / hour): Total bias-corrected precipitation, over land only. PRECTOTLAND variable in MERRA-2, converted from kg m⁻² s⁻¹

Snowfall (mm / hour): Total bias-corrected precipitation in the form of snow, over land only. PRECSNOLAND variable in MERRA-2, converted from kg m⁻² s⁻¹

Snow mass (kg / m²): Amount of snow per land area. SNOMAS variable in MERRA-2, in native units.

Air density (kg / m³): Air density at ground level. RHOA variable in MERRA-2, in native units.

Ground-level solar irradiance (W / m²): Surface-level incident shortwave radiation flux, considering cloud cover and aerosols. SWGDN variable in MERRA-2, in native units. Note, MERRA-2 uses the mean annual climate for aerosols, rather than time-varying quantity.

Top of atmosphere solar irradiance (W / m²): Top of atmosphere incident shortwave radiation flux, before cloud cover and aerosol influences. SWTDN variable in MERRA-2, in native units.

Cloud cover fraction: Fraction of cloud cover, averaged over grid cell and summed over all height above ground. CLDTOT variable in MERRA-2, in native units (a [0, 1] scale).


Check for multicollinearity among selected variables (Precipitation, IrradianceSurface, IrradianceToa, Snowfall, SnowMass, CloudCover, AirDensity).


```{r}
## VIF (Variance Inflation Factor)

library(car)

model.VIF <- lm(Temperature ~ Precipitation + IrradianceSurface + IrradianceToa + Snowfall + SnowMass + CloudCover + AirDensity, data = Ontario.Daily)

summary(model.VIF)

# Calculate VIF for selected variables
vif_values <- vif(model.VIF)

# Print VIF values
print(vif_values)
```


The Variance Inflation Factor (VIF) is a measure that quantifies how much the variance of an estimated regression coefficient increases if your predictors are correlated. High VIF values (typically > 10) may indicate problematic levels of multicollinearity.

Only for IrradianceSurface we observe a VIF value greater than 10 and VIF of IrradianceToa is close to 10.


```{r}
## Correlation Matrix

Temp.Ontario.Daily <- Ontario.Daily

colnames(Temp.Ontario.Daily) <- c("Date","Demand","Temperature","Holiday",
                                  "Precipitation","Irradiance (surface)",
                                  "Irradiance (atmosphere)","Snowfall",
                                  "Snow Mass","Cloud Cover","Air Density",
                                  "Day_Type","Temperature2",
                                  "HolidayNew","Day_TypeNew")

# Create a correlation matrix for selected variables
cor_matrix <- cor(Temp.Ontario.Daily[, c("Precipitation", "Irradiance (surface)", "Irradiance (atmosphere)", "Snowfall",  "Snow Mass",  "Cloud Cover", "Air Density")])

# Print the correlation matrix
print(cor_matrix)
```


High correlation coefficients (close to 1 or -1) may suggest multicollinearity. 

Corr(IrradianceToa, IrradianceSurface) = 0.8585024
Corr(IrradianceSurface, CloudCover) = -0.6143980
Corr(IrradianceSurface, AirDensity) = -0.6453389
Corr(IrradianceToa, AirDensity) = -0.7476595


We can visualize the correlation matrix using tools like corrplot for better insight:

```{r}
library(corrplot)

# Plot the correlation matrix
corrplot(cor_matrix, method = "color", tl.col="black")
```


Thus, we may omit "IrradianceSurface" and "IrradianceToa" from future models for Temperature to avoid multicollinearity.


### Choose best model to obtain fuzzily forecast of Temperature


Note that some models consider only lagged values of temperature. However, in some models we incorporate other covariates as well.


```{r}
### Mean Model
MeanModel.Temp <- Ontario.Daily %>%
  model(MEAN(Temperature))

fore.Mean.Temp <- MeanModel.Temp %>% forecast(Ontario.Daily)

Mean.Test.Accuracy <- fabletools::accuracy(fore.Mean.Temp, Ontario.Daily)

xtable(Mean.Test.Accuracy)
```



```{r}
## Naive Model
NaiveModel.Temp <- Ontario.Daily %>%
  model(NAIVE(Temperature))

fore.Naive.Temp <- NaiveModel.Temp %>% forecast(Ontario.Daily)

Naive.Test.Accuracy <- fabletools::accuracy(fore.Naive.Temp, Ontario.Daily)

xtable(Naive.Test.Accuracy)
```



```{r}
### Seasonal Naive Model
SNaiveModel.Temp <- Ontario.Daily %>%
  model(SNAIVE(Temperature))

fore.SNaive.Temp <- SNaiveModel.Temp %>% forecast(Ontario.Daily)

SNaive.Test.Accuracy <- fabletools::accuracy(fore.SNaive.Temp, Ontario.Daily)

xtable(SNaive.Test.Accuracy)
```



```{r}
### Drift Model
DriftModel.Temp <- Ontario.Daily %>%
  model(RW(Temperature~drift()))

fore.Drift.Temp <- DriftModel.Temp %>% forecast(Ontario.Daily)

Drif.Test.Accuracy <- fabletools::accuracy(fore.Drift.Temp, Ontario.Daily)

xtable(Drif.Test.Accuracy)
```



```{r}
library(fable.prophet)

### Prophet Model
ProphetModel.Temp1 <- Ontario.Daily %>%
  model(prophet(Temperature))

fore.Prophet.Temp1 <- ProphetModel.Temp1 %>% forecast(Ontario.Daily)

Prophet.Test.Accuracy1 <- fabletools::accuracy(fore.Prophet.Temp1, Ontario.Daily)

xtable(Prophet.Test.Accuracy1)
```



```{r}
### Prophet Model
ProphetModel.Temp2 <- Ontario.Daily %>%
  model(prophet(Temperature ~ Precipitation + Snowfall + SnowMass + CloudCover + AirDensity))

fore.Prophet.Temp2 <- ProphetModel.Temp2 %>% forecast(Ontario.Daily)

Prophet.Test.Accuracy2 <- fabletools::accuracy(fore.Prophet.Temp2, Ontario.Daily)

xtable(Prophet.Test.Accuracy2)
```



```{r}
### ARIMA Model
ARIMAModel.Temp1 <- Ontario.Daily %>%
  model(ARIMA(Temperature))

## Note: Here manipulate date variable to obtain accuracy. 
## Forecasts from an ARIMA model must start one step beyond the end of the trained data.
test <- Ontario.Daily %>%
  mutate(Date = seq(as.Date("2020-01-01"), as.Date("2021-12-30"), by = "day"))
  
fore.ARIMA.Temp1 <- ARIMAModel.Temp1 %>% forecast(test)

ARIMA.Test.Accuracy1 <- fabletools::accuracy(fore.ARIMA.Temp1, test)

xtable(ARIMA.Test.Accuracy1)
```



```{r}
### ARIMA Model
ARIMAModel.Temp2 <- Ontario.Daily %>%
  model(ARIMA(Temperature ~ Precipitation + Snowfall + SnowMass + CloudCover + AirDensity))

## Note: Here manipulate date variable to obtain accuracy. 
## Forecasts from an ARIMA model must start one step beyond the end of the trained data.
test <- Ontario.Daily %>%
  mutate(Date = seq(as.Date("2020-01-01"), as.Date("2021-12-30"), by = "day"))
  
fore.ARIMA.Temp2 <- ARIMAModel.Temp2 %>% forecast(test)

ARIMA.Test.Accuracy2 <- fabletools::accuracy(fore.ARIMA.Temp2, test)

xtable(ARIMA.Test.Accuracy2)
```



```{r}
### nnetar Model
nnetarModel.Temp1 <- Ontario.Daily %>%
  model(NNETAR(Temperature))

fore.nnetar.Temp1 <- nnetarModel.Temp1 %>% forecast(Ontario.Daily)

nnetar.Test.Accuracy1 <- fabletools::accuracy(fore.nnetar.Temp1, Ontario.Daily)

xtable(nnetar.Test.Accuracy1)
```



```{r}
### nnetar Model
nnetarModel.Temp2 <- Ontario.Daily %>%
  model(NNETAR(Temperature ~ Precipitation + Snowfall + SnowMass + CloudCover + AirDensity))

fore.nnetar.Temp2 <- nnetarModel.Temp2 %>% forecast(Ontario.Daily)

nnetar.Test.Accuracy2 <- fabletools::accuracy(fore.nnetar.Temp2, Ontario.Daily)

xtable(nnetar.Test.Accuracy2)
```



```{r}
### EWMA Model
EWMAModel.Temp <- Ontario.Daily %>%
  model(ETS(Temperature))

fore.EWMA.Temp <- EWMAModel.Temp %>% forecast(Ontario.Daily)

EWMA.Test.Accuracy <- fabletools::accuracy(fore.EWMA.Temp, Ontario.Daily)

xtable(EWMA.Test.Accuracy)
```



```{r}
## TSLM
TSLMModel.Temp1 <- Ontario.Daily %>%
  model(TSLM(Temperature))

fore.TSLM.Temp1 <- TSLMModel.Temp1 %>% forecast(Ontario.Daily)

TSLM.Test.Accuracy1 <- fabletools::accuracy(fore.TSLM.Temp1, Ontario.Daily)

xtable(TSLM.Test.Accuracy1)
```



```{r}
## TSLM

TSLMModel.Temp2 <- Ontario.Daily %>%
  model(TSLM(Temperature ~ Precipitation + Snowfall + SnowMass + CloudCover + AirDensity))

fore.TSLM.Temp2 <- TSLMModel.Temp2 %>% forecast(Ontario.Daily)

TSLM.Test.Accuracy2 <- fabletools::accuracy(fore.TSLM.Temp2, Ontario.Daily)

xtable(TSLM.Test.Accuracy2)
```



```{r}
Temp.Model.Accuracy <- rbind(Mean.Test.Accuracy, Naive.Test.Accuracy, SNaive.Test.Accuracy, Drif.Test.Accuracy, Prophet.Test.Accuracy1, Prophet.Test.Accuracy2, ARIMA.Test.Accuracy1, ARIMA.Test.Accuracy2, nnetar.Test.Accuracy1, nnetar.Test.Accuracy2, EWMA.Test.Accuracy, TSLM.Test.Accuracy1, TSLM.Test.Accuracy2)
Temp.Model.Accuracy
```



Note that considering all the models listed in the table, nnetar (Temperature ~ Precipitation + Snowfall + SnowMass + CloudCover + AirDensity) has the lowest errors. Thus, this model will be used to obtain fuzzy forecasts of temperature.


Forecast plot for the best fit model for Temperature:


```{r}
t <- seq(1:nrow(Ontario.Daily))
plot(t, Ontario.Daily$Temperature, col="blue", xlab = "Day", ylab ="Temperature (°C)", type="l", lwd = 2, ylim = c(-32,30))
lines(t, fore.nnetar.Temp2$.mean, col="red", lwd = 2)
legend("bottomleft", legend = c("Observed", "Forecasts"), col = c("blue","red"), lty = 1, lwd = 2)
```


## New Plot

```{r}
nn.Temp.table <- tsibble(
  Date = Temp.Ontario.Daily$Date,
  Observed = Temp.Ontario.Daily$Temperature,
  Forecasts = fore.nnetar.Temp2$.mean
)

library(ggplot2)

ggplot(data = nn.Temp.table, aes(x = Date)) +
  geom_line(aes(y = Observed, color = "Observed")) +
  geom_line(aes(y = Forecasts, color = "Forecasts")) +
  labs(y = "Temperature (°C)",
       x = "Date") +
  scale_color_manual(name = "Data",
                     values = c(Observed = "blue", Forecasts = "red"),
                     labels = c(Observed = "Observed", Forecasts = "Forecasts")) +
  theme_minimal()
```





### Fit different models to identify the best fit model for electricity demand


```{r}
## Train and testing datasets
## Training data: 80% of data 730*0.80 = 584
## Testing data: 730-584 = 146

trainOntario = Ontario.Daily[1:584,]
testOntario =  Ontario.Daily[585:730,]
head(trainOntario)
head(testOntario)
```



```{r}
## TSLM
TSLM.Reg.Demand <- trainOntario %>%
  model(tslm = TSLM(Demand ~ Temperature + Temperature2 + HolidayNew + Day_TypeNew))

report(TSLM.Reg.Demand)

TSLM.Reg.fore <- TSLM.Reg.Demand %>% forecast(testOntario)

TSLM.Reg.Demand.Accuracy <- fabletools::accuracy(TSLM.Reg.fore, testOntario)

xtable(TSLM.Reg.Demand.Accuracy)

```



```{r}
## ARIMA model
ARIMA.Reg.Demand <- trainOntario %>%
  model(arima =  ARIMA(Demand ~ Temperature + Temperature2 + HolidayNew + Day_TypeNew))

report(ARIMA.Reg.Demand)

ARIMA.Reg.fore <- ARIMA.Reg.Demand %>% forecast(testOntario)

ARIMA.Reg.Demand.Accuracy <- fabletools::accuracy(ARIMA.Reg.fore, testOntario)

xtable(ARIMA.Reg.Demand.Accuracy)

```




```{r}
## nnetar model
nnetar.Reg.Demand <- trainOntario %>%
  model(NNETAR(Demand ~ Temperature + Temperature2 + HolidayNew + Day_TypeNew))

nnetar.Reg.Demand

nnetar.Reg.fore <- nnetar.Reg.Demand %>% forecast(testOntario)

nnetar.Reg.Demand.Accuracy <- fabletools::accuracy(nnetar.Reg.fore, testOntario)

xtable(nnetar.Reg.Demand.Accuracy)
```



```{r}
### Prophet Model
Prophet.Reg.Demand <- trainOntario %>%
  model(prophet(Demand ~ Temperature + Temperature2 + HolidayNew + Day_TypeNew))

Prophet.Reg.Demand %>% components() %>% autoplot()

Prophet.Reg.fore <- Prophet.Reg.Demand %>% forecast(testOntario)

Prophet.Reg.Demand.Accuracy <- fabletools::accuracy(Prophet.Reg.fore, testOntario)

xtable(Prophet.Reg.Demand.Accuracy)
```



```{r}
Demand.Model.Accuracy <- rbind(TSLM.Reg.Demand.Accuracy, ARIMA.Reg.Demand.Accuracy, nnetar.Reg.Demand.Accuracy, Prophet.Reg.Demand.Accuracy)
Demand.Model.Accuracy
```


Note that among the 4 regression models consider for Demand, nnetar has the best forecast accuracy. Thus, nnetar will be used to obtain fuzzy forecasts of Demand using fuzzy forecasts of Temperature.



Forecast plot for the best fit model for Demand:


```{r}
t <- seq(1:nrow(testOntario))
plot(t, testOntario$Demand, col="blue", xlab = "Day", ylab ="Electricity Demand (MW)", type="l", lwd = 2, ylim = c(10500, 19000))
lines(t, nnetar.Reg.fore$.mean, col="red", lwd = 2)
legend("bottomleft", legend = c("Observed", "Forecasts"), col = c("blue","red"), lty = 1, lwd = 2)
```




## New Plot

```{r}
nn.Temp.table <- tsibble(
  Date = testOntario$Date,
  Observed = testOntario$Demand,
  Forecasts = nnetar.Reg.fore$.mean
)

ggplot(data = nn.Temp.table, aes(x = Date)) +
  geom_line(aes(y = Observed, color = "Observed")) +
  geom_line(aes(y = Forecasts, color = "Forecasts")) +
  labs(y = "Electricity Demand (MW)",
       x = "Date") +
  scale_color_manual(name = "Data",
                     values = c(Observed = "blue", Forecasts = "red"),
                     labels = c(Observed = "Observed", Forecasts = "Forecasts")) +
  scale_x_date(date_labels = "%Y-%m-%d", date_breaks = "1 month") +  # Specify date format and breaks
  theme_minimal()
```



